Reproducible Research - June 2014
========================================================
Ron Mashrouteh - Reproducible Research - Peer assignment 1

loading downloaded file **(activity.csv)**

```{r echo=TRUE}
ds <- read.csv("~/activity.csv")
```

creating my dataset (myds) to get setps, dates and intervals without NA

```{r echo=TRUE}
myds <- subset(ds, na.rm=TRUE, steps!=0, select=c(steps, date, interval))
```

globally declare **myds** so I don't have to keep typing myds$....

```{r echo=TRUE}
attach(myds)
```

mean total number of steps taken

```{r echo=TRUE}
mean(ds$steps, na.rm=TRUE)
```

histogram graph - steps taken per date

```{r echo=TRUE}
hist(steps, breaks=100, main="Histogram Graph - Steps", ylab="Date", xlab="steps")
```

**mean** and **median** total number of steps per day

```{r echo=TRUE}
perdaymean <- aggregate(list(steps=steps), na.rm=TRUE , list(dates = date), mean)
perdaymean
perdaymedian <- aggregate(list(steps=steps), na.rm=TRUE , list(date = date), median)
perdaymedian
```

**Start preparing for plotting a graph of 5 minute interval**

minimum number of steps

```{r echo=TRUE}
mi <- min(interval)
mi
```

maximum number of steps

```{r echo=TRUE}
mx <- max(interval)
mx
```

Generating aggreate of average steps per interval

```{r echo=TRUE}
perdayavg <- aggregate(list(steps = steps), na.rm=TRUE , list(interval = interval), mean)
perdayavg
```

Plotting the graph according to the aggregate we generated above

```{r echo=TRUE}
plot(perdayavg, type = "l", xlim=c(mi, mx), lwd = 2, main="Time Interval", ylab="Avg. steps all day", xlab="5-minute interval", col="blue")
```

**which 5-minute interval contains the maximum number of steps**

MaxVal_avgsteps - get the maximum number of steps
InterVal - get the Interval value for MaxVal_avgsteps.
```{r echo=TRUE}
MaxVal_avgsteps <- max(perdayavg$steps)
InterVal <- subset(perdayavg$interval, perdayavg$steps == MaxVal_avgsteps)

MaxVal_avgsteps
InterVal
```

Inputing missing value
========================================================

Loading the downloaded file as a new dataset

```{r echo=TRUE}
Nwds <- read.csv("~/activity.csv")
```

Calculating total number of missing values (NAs)

```{r echo=TRUE}
NAs <- length(Nwds$step[Nwds$step =="NA"])
NAs
```

Subset & Fill the missing value with 0s (replacing NA with 0)

```{r echo=TRUE}
Nwds <- subset(Nwds, select=c(steps, date, interval))
Nwds
```

Setting to 0s

```{r echo=TRUE}
Nwds[is.na(Nwds)] <- 0
Nwds
```

Remove all the rows that contain NA values

We load a fresh copy of the csv file to get all data with NA since we substitute them with 0s in the previous step

```{r echo=TRUE}
Newds <- read.csv("~/activity.csv")
Newds
```

Remove all rows that contan NA

```{r echo=TRUE}
remv <-na.omit(Newds)
remv
```

Calculate the mean and median values after removing all rows that contained NA

```{r echo=TRUE}
perdayavg_mean_noNA <- aggregate(list(steps = remv$steps), list(interval = remv$interval), mean)
perdayavg_median_noNA <- aggregate(list(steps = remv$steps), list(interval = remv$interval), median)
```

Generate histogram graphs for mean and median

```{r echo=TRUE}
hist(perdayavg_mean_noNA$steps, breaks=100, main="Histogram Graph After NA removal- Steps", ylab="Interval", xlab="steps")
```

```{r echo=TRUE}
hist(perdayavg_median_noNA$steps, breaks=10, main="Histogram Graph After NA removal- Steps", ylab="Interval", xlab="steps")
```

Loadin the csv file one more time for the last part of the assignment

```{r echo=TRUE}
lastpart <- read.csv("~/activity.csv")

NAs <- length(lastpart$step[lastpart$step =="NA"])
shwval <- NAs

lastpart[is.na(lastpart)] <- 0
```

Specifying the if the date is a weekday or weekend

```{r echo=TRUE}
lastpart$weekday = weekdays(as.Date(lastpart$date, format = "%Y-%m-%d"))
lastpart$weekday.type = factor(ifelse(lastpart$weekday == "Sunday" | lastpart$weekday == 
                                             "Saturday", "weekend", "weekday"), levels = c("weekday", "weekend"))


lastpart.weekdays = lastpart[lastpart$weekday.type == "weekday", ]
lastpart.weekend = lastpart[lastpart$weekday.type == "weekend", ]
avg.steps.per.interval.weekdays = sapply(split(lastpart.weekdays$steps, lastpart.weekdays$interval), mean)
avg.steps.per.interval.weekend = sapply(split(lastpart.weekend$steps, lastpart.weekend$interval), mean)
```

Creating the panel plot for the 5-minute interval.

```{r echo=TRUE}
DoPlot = function() {
    par(mfrow = c(2, 1), mar = c(4, 5, 2, 2))
    plot(avg.steps.per.interval.weekend, type = "l", yaxt = "n", ylim = c(0, 250), xlim = c(0, 300), main = "weekend", ylab = "", xlab = "Interval")
    axis(side = 4, at = seq(0, 250, 50), labels = seq(0, 250, 50))
    plot(avg.steps.per.interval.weekdays, type = "l", ylim = c(0, 250), xlim = c(0, 300), main = "weekday", ylab = "", xlab = "Interval")
    par(mfrow = c(1, 1), mar = c(3, 2, 2, 2))
    mtext(text = "Number of steps", side = 2)
    par(mar = c(5, 5, 5, 2))
}

DoPlot()
```

